---
title: "Holt-Winters"
output: html_notebook
fig_width: 12 
fig_height: 4 
---

This builds an example of the Holt-Winters forecasting method to look for anomalies.


```{r setup, echo=FALSE, message=FALSE}

# # Get Google Analytics credentials and set them. 
# ga_client_id <- Sys.getenv("GA_CLIENT_ID")
# ga_client_secret <- Sys.getenv("GA_CLIENT_SECRET")
# 
# options(googleAuthR.client_id = ga_client_id)
# options(googleAuthR.client_secret = ga_client_secret) 

packages <- c("tidyverse","googleAnalyticsR", "splitstackshape", "knitr", "kableExtra", "extrafont", "tools", "scales")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}

library(tidyverse)
# library(googleAnalyticsR)    # Uncomment if actually pulling data
library(splitstackshape)
library(knitr)
library(kableExtra)
library(extrafont)
library(tools)
library(scales)
library(lubridate)

# Get the themes
source("scripts/themes_nunito.R")

# Set the number of time periods to be checked. These will be the "last X rows" in
# the data that gets pulled
check_periods <- 7

# Set the frequency of the time series. If dimensions is "date," then this should be 7.
# This doesn't currently work for anything other than daily data
ts_freq <- 7

# Set the level for the prediction interval
interval_level <- 0.95

#################
# Updates got tedious, so we're just going to generate a faux data set that looks plausible. 
# The code for actually pulling the data from GA is still included below should you want
# to do this with real data.
# 
# end_date <- as.Date("2018-10-27")
# start_date <- end_date - 7*7+1    # Pull 7 weeks of data
# 
# date_range <- c(start_date, end_date)
# 
# # Authorize Google Analytics
# ga_auth()
# 
# # Get the view ID to be used
# view_id <- Sys.getenv("GA_VIEW_ID")
# 
# metric <- "sessions"
# dimension <- "date"  # Set this to day or week. This hasn't been vetted for anything other than "date"

# # Get the data
# ga_data <- google_analytics(view_id,
#                               date_range = c(start_date, end_date),
#                               dimensions = dimension,
#                               metrics = metric)
# 
# # Rename columns
# names(ga_data) <- c("time_period", metric)

# saveRDS(ga_data, "ga_data.rds")
# ga_data <- readRDS("ga_data.rds")
#################


#############
# Generate a Data Set -- 49 days
#############

# Set seed for reproducibility
set.seed(617)

# Set up a data frame with a recurring weekly pattner
ga_data <- data.frame(time_period = seq(as.Date("2018-09-09"), as.Date("2018-10-27"), by = "day"),
                      sessions = rep(c(5000, 22750, 24000, 24250, 24000, 23000, 5750),7))

# Add some noise to the data
ga_data$sessions <- ga_data$sessions + runif(49, 1, 1500) %>% round()

# Decline for the first two weeks, trend up for the next 4 weeks, fudge for the last week
trend_adjust <- c(seq(from = 0, by = -100, length.out = 14),
                  seq(from = -1400, by = 125, length.out = 28),
                  2000, 0, 1500, 1700, 1500, 700, 2000)

# Add the trend to the data
ga_data$sessions <- ga_data$sessions + trend_adjust



# Determine how many rows of the data will be used to build the forecast. This
# is just ALL rows minus the number of periods to be checked
rowcount_forecast <- nrow(ga_data) - check_periods

# Also figure out the date where the cutoff is between training and forecast
cutoff_time_period <- ga_data[rowcount_forecast,1] 

# We actually want to shift this over a little bit to fall between two points when we plot
cutoff_time_period <- (2*as.numeric(cutoff_time_period) + 1)/2

# Make a data set that removes the "rows to be evaluated." This will get 
# used both to generate the time series for the forecast as well as for modeling
ga_data_training <- ga_data %>%
  top_n(-rowcount_forecast, wt = time_period) 

# Get the time_period values for the forecast period
time_periods_forecast <- ga_data %>%
    top_n(check_periods, wt = time_period) %>%
  select(time_period)

# Make a time-series object using the data for the training period. This
# is what we'll use to build the forecast
ga_data_ts <- ga_data_training[[2]] %>%
  ts(frequency = ts_freq)

# Start building out our master data for plotting by adding a column that
# has just the data being used for the training
ga_data_plot <- ga_data %>%
  left_join(ga_data_training, by = c(time_period = "time_period"))

# Rename to generic names
names(ga_data_plot) <- c("time_period", "metric_all", "metric_training")

# Add a column that is just the actuals data of interest
ga_data_plot <- ga_data_plot %>%
  mutate(metric_assess = ifelse(is.na(metric_training), metric_all, NA))

# Generate a Holt Winters forecast
hw <- HoltWinters(ga_data_ts)

# Predict the next 7 days (the 7 days of interest). Go ahead and convert it to a data frame
forecast_metric <- predict(hw, n.ahead = 7, prediction.interval = T, interval_level = 0.95) %>%
  as.data.frame()

# Add in the dates so we can join this with the original data. We know it was the 7 days
# starting from cutoff_date
forecast_metric$time_period <- time_periods_forecast$time_period

# Add these columns to the original data and add a column that IDs anomaly points by 
# checking to see if the actual value is outside the upper or lower bounds. If it is,
# If it is, put the value. We'll use this to highlight the anomalies.
ga_data_plot <- ga_data_plot %>%
  left_join(forecast_metric) %>%
  mutate(anomaly = ifelse(metric_all < lwr | metric_all > upr, metric_all, NA))

# Figure out the max value (for plotting) and then bump it up a bit
max_y <- max(select(ga_data_plot, -time_period))

```

## And...Usually...A Trend of the Data

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Make a weekly chart of the data

# Aggregate the data to be weekly
ga_data_weekly <- ga_data %>%
  mutate(week = time_period - wday(time_period) + 1) %>% 
  group_by(week) %>% 
  summarise(sessions = sum(sessions)) %>% 
  mutate(last_week_sessions = ifelse(week == max(week), sessions, NA ))

# Chart it
ga_plot_weekly <- ggplot(ga_data_weekly, mapping = aes(x = week, y = sessions)) +
  geom_line(color = "#0060AF", size = 1) +
  geom_point(aes(y=sessions), size=3, shape = 19, colour = "#0060AF") +
  geom_point(aes(y=sessions), size=3, shape = 20, colour = "white") +
  geom_point(aes(y=last_week_sessions), size=5, colour = "white") +
  geom_point(aes(y=last_week_sessions), size=3.5, colour = "#9A258F") +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(0, 150000)) +
  labs(x = " ") +
  theme_hw

ga_plot_weekly

```

## Let's break that weekly data down to be daily.

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Build a plot showing just the actual data
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_line(aes(y = ga_data_plot$metric_all), color = "#0060AF", size = 0.75) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(0, 30000)) +
  labs(x= " ") +
  theme_hw

ga_plot

```

## Last week is now the last 7 data points.

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, but showing just the last week
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_line(aes(y = ga_data_plot$metric_all), color = "#0060AF", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(0, 30000)) +
  labs(x= " ") +
  theme_hw

ga_plot

```

## We have the data of interest...

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, with the "training data highlighted"data of interest highlighted
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_line(aes(y = ga_data_plot$metric_training), color = "gray80", size = 0.75) +
  geom_line(aes(y = ga_data_plot$metric_assess), color = "#0060AF", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(0, 30000)) +
  labs(x= " ") +
  theme_hw

ga_plot

```

## ...and data for context.

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, with the training data highlighted
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_line(aes(y = ga_data_plot$metric_training), color = "#0060AF", size = 0.75) +
  geom_line(aes(y = ga_data_plot$metric_assess), color = "gray80", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(0, 30000)) +
  labs(x= " ") +
  theme_hw

ga_plot

```

## Let's take our context data...

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, with the training data highlighted
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_line(aes(y = ga_data_plot$metric_training), color = "#0060AF", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(0, 30000)) +
  labs(x= " ") +
  theme_hw

ga_plot

```

## ...and think of it as "training" data.

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, with the training data highlighted
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_line(aes(y = ga_data_plot$metric_training), color = "#0060AF", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(0, 30000)) +
  labs(x= " ") +
  theme_hw

ga_plot

# Do another version without the horizontal gridlines
ga_plot +
  theme(panel.grid.major.y = element_blank())

```

## We can "decompose" that data.

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Decompose the time-series data
ga_decomp <- decompose(ga_data_ts) 

# Convert that to a long format data frame
ga_decomp_df <- data.frame(
  Actual = ga_decomp$x,
  Seasonal = ga_decomp$seasonal,
  Trend = ga_decomp$trend,
  Random = ga_decomp$random) %>%
  mutate(time_period = ga_data_training$time_period) %>%
  gather(key, value, -time_period)

# We want to control the order of the output, so make key a factor
ga_decomp_df$key <- factor(ga_decomp_df$key,
                           levels = c("Actual", "Seasonal", "Trend", "Random"))

## We can "decompose" that data.

# The final plot will show all of the data.
ga_decomp_df_plot <- ga_decomp_df 
  
# Plot the values
ga_plot <- ggplot(ga_decomp_df_plot, mapping = aes(x = time_period, y = value, colour = key)) +
  geom_line(size = 1) +
  facet_grid(key ~ ., scales = "free", switch = "y") +
  scale_color_manual(values=c("#0060AF", "#999999", "#999999", "#999999")) +
  theme_sparklines

ga_plot

```

## We can now build our forecast.

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, with the training data highlighted
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_line(aes(y = ga_data_plot$metric_training), color = "#0060AF", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  geom_line(aes(y = ga_data_plot$fit), color = "gray50", linetype = "dotted", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(-1000, max_y)) +
  theme_hw

ga_plot

```

## The forecast won't be perfect.

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, with the forecasted value shown
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_line(aes(y = ga_data_plot$metric_training), color = "#0060AF", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  geom_line(aes(y = ga_data_plot$fit), color = "gray50", linetype = "dotted", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(-1000, max_y)) +
  theme_hw

ga_plot

```

## So we add a prediction interval

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, with the prediction interval added
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_ribbon(aes(ymin = ga_data_plot$lwr, ymax = ga_data_plot$upr), fill = "gray90") +
  geom_line(aes(y = ga_data_plot$metric_training), color = "#0060AF", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  geom_line(aes(y = ga_data_plot$fit), color = "gray50", linetype = "dotted", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(-1000, max_y)) +
  theme_hw

ga_plot

```

## We can compare!

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

# Same plot, with the actuals shown
ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_ribbon(aes(ymin = ga_data_plot$lwr, ymax = ga_data_plot$upr), fill = "gray90") +
  geom_line(aes(y = ga_data_plot$metric_all), color = "#0060AF", size = 0.75) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  geom_line(aes(y = ga_data_plot$fit), color = "gray50", linetype = "dotted", size = 1) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(-1000, max_y)) +
  theme_hw

ga_plot

```

## Now we have meaningful context! 

```{r echo=FALSE, message=FALSE, warning = FALSE, fig.height=5, fig.width=10}

ga_plot <- ggplot(ga_data_plot, mapping = aes(x = time_period)) +
  geom_ribbon(aes(ymin = ga_data_plot$lwr, ymax = ga_data_plot$upr), fill = "gray90") +
  geom_line(aes(y = ga_data_plot$metric_all), color = "#0060AF", size = 0.75) +
  geom_line(aes(y = ga_data_plot$fit), color = "gray50", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = cutoff_time_period), 
             color = "gray40", linetype = "dashed", size = 1) +
  geom_point(aes(y = ga_data_plot$anomaly), color = "#F58220", size = 2.5) +
  scale_y_continuous(label=comma, expand = c(0, 0), limits = c(-1000, max_y)) +
  theme_hw

ga_plot

```
